Automated Topic Naming

### 2012-01-17 ese/ese.tex
fixes for review comments, including new process model diagram.

\section{Introduction}
A key problem for practicing software maintainers is gaining an
understanding of \emph{why} a system has evolved the way it has \cite{Mockus00}. 
This is different from \emph{how} a system has evolved. 
% NE this phrase confuses me: "because the change of behaviour itself is the \emph{how}.
Looking back on streams of artifacts scattered across different
repositories, inferring what activities were performed, when, and for
what reasons, is hard without expert advice from the developers
involved. 
%new stuff
% edit this. awkward
In this work we seek to provide a method of automatically labelling  development topics extracted from commit logs.
Concrete applications of topics and \emph{labelled topic extraction}, which this paper introduces,
range from project dashboards to annotating software artifacts such as revisions and bug reports with
NFR-related tags.
Project dashboards~\cite{dashboard} are typically employed by managers and are used to
provide quick summaries of the effort put into a software project. In this case, 
labelled topic extraction would allow managers to track effort
related to NFR topics, such as \emph{portability}.
These techniques also allow for the annotation of commit comments
 and other software artifacts with NFRs. 
This would enable querying of bug reports and artifacts by relevant NFRs. 
For instance a manager can confirm if their developers were
focused on \emph{usability} by looking for
 \emph{usability}-relevant revisions and bug reports.

@@

In this section we describe how to label topics based on dictionaries
mined from sources external to the projects.

@@

In order to automatically label each topic with 
one of the six high-level NFRs,
%an NFR from ISO9126, 
we associate each NFR with a list
of keywords (\emph{word-lists}, in our parlance). These word-lists were determined a priori and were not extracted from the projects themselves.
We intersected the words of the topics and the words of our word-lists.
We ``labelled'' a topic if any of its words matched any of the word-list's words.
A topic could match more than one NFR.
We used several different sets of word-lists for comparison, which we
refer to as \textsf{exp1}, \textsf{exp2}, and \textsf{exp3} in the text which follows.

Our first word-list set, \textsf{exp1}, was generated using the ontology described in Kayed et al.~\cite{5072519}
That paper constructs an ontology for software quality measurement using eighty source documents, including research papers and international standards. 
The labels we used:

@@

%\DONE{this discussion is about which DB? where do the numbers come from?}
For \textsf{exp1} the labels with the most topics
were
\emph{correctness} (182/305/748 -- MySQL, MaxDB and PostgreSQL topics) and \emph{testability} (121/238/625). 
We did not see many results for \emph{usability} (4/0/138) or
\emph{accuracy} (3/0/27), which were infrequently matched. Note the significantly higher result for \emph{usability} in PostgreSQL -- this suggests a difference in how this project is discussing usability, at least with respect to our analysis.
We also looked for correlations between our labels: excluding double
matches (self-correlation), our highest co-occurring terms were
\emph{verifiability} or \emph{correctness}
with \emph{traceability}, and \emph{testability} with \emph{correctness}.% (76 and 62 matches, respectively).
% See the following sections for our error analysis.

@@

\subsubsection{Analysis of the Semi-unsupervised Labelling} %--- %Based
				%on the labels, and our manual topic
				%labelling, 
%\noindent \textbf{Analysis of the unsupervised labelling} --- %Based on the labels, and our manual topic labelling, 
%we compared the results of the unsupervised word matching approach. 
For each quality in the high-level ISO 9126 taxonomy (namely, \emph{Maintainability, Usability, Reliability, Efficiency, Portability, Functionality}) we assessed whether the semi-unsupervised labels for a topic matched the manual annotations we created for the validation corpus. 
To be clear, the manual annotations were not used to
train the labelling process.
As described in Section \ref{sec:wordlist} we used both ROC and F-1 measures to evaluate the performance of the classification.
Figure \ref{fig:maxdb-unsup-results} shows our ROC results for PostgreSQL, MaxDB and MySQL. We describe F-1 results in the text below.

@@

Because our ground truth annotations were relevant only to ISO9126,
we estimate that \textsf{exp1} had poor
performance via the overlap between ISO9126 and the Kayed ontology.
For \textsf{exp1} the F-measures for the NFRs for MaxDB were from $0$ to $0.18$ with an average (of all NFRs)
of $0.03$, for MySQL were from $0$ to $0.16$ with an average of
$0.05$, and for PostgreSQL $0$ to $0.15$ with an average of $0.07$.

@@

\subsubsection{Analysis of the Supervised Labelling}
% \label{sec:suplabelling}
Because our data-set was of word counts we expected Bayesian techniques, often used in spam filtering, to perform well. 
We tried other learners that WEKA~\cite{weka09} provides: rule learners, decision tree learners, vector space learners, and support vector machines.  
Figure \ref{fig:best-learn-per-tag} shows the performance of the best
performing learner per label: 
% We considered the best learner for a label to be 
the learner that had the highest ROC value for that label. 
The best learner is important because one uses a single learner per
label. When applying our technique, for each NFR one should select the
best performing learner possible.
%Figure \ref{fig:best-learn-per-tag} uses the ZeroR learner as a baseline, a common practice in machine learning,
%as ZeroR naively chooses the largest category all of the time. 
%ZeroR occasionally outperforms other learners; for labels which are not as common, this is to be expected because any miscategorization will hurt accuracy. 
%This is why we use ROC values, instead of accuracy, as they can better represent performance on labels which are not applicable to the majority of samples.

Figure \ref{fig:best-learn-per-tag} shows that MaxDB and MySQL have
quite different results, as the ROC values for \emph{reliability} and
\emph{functionality} swap between projects. For PostgreSQL, the
performance is nearly always more poor than the other two systems. The
reason for this lack of performance could be that the number of
topics, $N$ could be non-optimal for PostgreSQL. Perhaps topics were
getting too mixed and it was becoming hard to distinguish one topic
from the next. PostgreSQL's data-set was the largest of the 3 projects:
the XML data-dump that described the PostgreSQL topics that we
annotated was 8 times larger than MySQL and 1.7 times larger than MaxDB. These
size differences arise from the number of commits, the number of files
and the verbosity of the commit descriptions.

@@

As noted in Section \ref{sec:wordlist}, each topic in our data-set can be composed of zero, one, or more NFRs. 
For example, a commit message might address \textit{reliability} in the context of \textit{efficiency}, or make a \textit{maintainability} improvement
in the source code that related to \textit{usability}. 
However, traditional machine learning techniques, such as Naive Bayes, can only map topics to a single class. 
The Mulan~\cite{mulan} library encapsulates several different multi-label machine learners which can label elements with multiple labels.
Mulan also includes methods for determining the performance of such techniques.
% TOO DETAILED The problem framed in the learners above has changed; instead of looking at the precision and recall of applying one label, we rank
% multiple labels at once. We must check if the full subset of labels was applied, and then how much of that subset was applied.

@@

\emph{Construct validity} -- we used only commit messages rather than
mail or bug tracker messages. 
To extend further we would need matching repositories for each project.
%Each case study was BecauseSince we required a seamless record, it was not possible to obtain
%other data sources. 
Possibly they would have influenced our results, but there would be a degree of correlation between the corpora.
%Conversely, construct validity was addressed in our validation as we used the files, topics and revisions to annotate the commit messages. 
%We used mailing-lists to verify the purpose behind some events.
Our taxonomy for software NFRs is subject to dispute, but seems to be
generally accepted. Finally, there are exogenous sources, such as
in-person discussions, which we did not access, an omission as noted
in Aranda et al.~\cite{aranda09icse} %We did not show these results to the stakeholders.

@@

Our data and scripts are available at:

\url{http://softwareprocess.es/nomen/}

@@

%\begin{acknowledgements}
%\end{acknowledgements}

>>>>

\section{Introduction}
A key problem for practicing software maintainers is gaining an
understanding of \emph{why} a system has evolved the way it has \cite{Mockus00}. 
This is different from \emph{how} a system has evolved. 
% NE this phrase confuses me: "because the change of behaviour itself is the \emph{how}.
Looking back on streams of artifacts scattered across different
repositories, inferring what activities were performed, when, and for
what reasons, is hard without expert advice from the developers
involved. 
%new stuff
% edit this. awkward
In this work we provide a method of automatically labelling  development topics extracted from commit logs.

Concrete applications of topics and \emph{labelled topic extraction}, which this paper introduces,
range from project dashboards to annotating software artifacts such as revisions and bug reports with
NFR-related tags.
Project dashboards~\cite{dashboard} are typically employed by managers and are used to
provide quick summaries of the effort put into a software project. In this case, 
labelled topic extraction would allow managers to track effort
related to NFR topics, such as \emph{portability}.
These techniques also allow for the annotation of commit comments
 and other software artifacts with NFRs. 
This would enable querying of bug reports and artifacts by relevant NFRs. 
For instance a manager can confirm if their developers were
focused on \emph{usability} by looking for
 \emph{usability}-relevant revisions and bug reports.

@@

In this section we describe how to label topics based on dictionaries
mined from sources external to the projects. We call this `semi-unsupervised' because while there is no ``training set'', we do seed the wordlists manually.

@@

In order to automatically label each topic with 
one of the six high-level NFRs,
%an NFR from ISO9126, 
we associate each NFR with a list
of keywords (\emph{word-lists}, in our parlance). These word-lists were determined a priori and were not extracted from the projects themselves, using the methodology explained below. In general, these lists are project and domain independent.
We intersected the words of the topics and the words of our word-lists.
We ``labelled'' a topic if any of its words matched any of the word-list's words.
A topic could match more than one NFR.
We used several different sets of word-lists for comparison, which we
refer to as \textsf{exp1}, \textsf{exp2}, and \textsf{exp3} in the text which follows.

Our first word-list set, \textsf{exp1}, was generated using the ontology described in Kayed et al.~\cite{5072519}
That paper constructs an ontology for software quality measurement using eighty source documents, including research papers and international standards. 
The labels we used are:

@@

%\DONE{this discussion is about which DB? where do the numbers come from?}
For \textsf{exp1} the labels with the most topics
were
\emph{correctness} (182/305/748, which represent MySQL, MaxDB and PostgreSQL topic counts, respectively) and \emph{testability} (121/238/625). 
We did not see many results for \emph{usability} (4/0/138) or
\emph{accuracy} (3/0/27), which were infrequently matched. Note the significantly higher result for \emph{usability} in PostgreSQL -- this suggests a difference in how this project is discussing usability, at least with respect to our analysis.
We also looked for correlations between our labels: excluding double
matches (self-correlation), our highest co-occurring terms were
\emph{verifiability} or \emph{correctness}
with \emph{traceability}, and \emph{testability} with \emph{correctness}.% (76 and 62 matches, respectively).
% See the following sections for our error analysis.

@@

\subsubsection{Analysis of the Semi-unsupervised Labelling} %--- %Based
				%on the labels, and our manual topic
				%labelling, 
%\noindent \textbf{Analysis of the unsupervised labelling} --- %Based on the labels, and our manual topic labelling, 
%we compared the results of the unsupervised word matching approach. 
For each quality in the high-level ISO 9126 taxonomy (namely, \emph{Maintainability, Usability, Reliability, Efficiency, Portability, Functionality}) we assessed whether the semi-unsupervised labels for a topic matched the manual annotations we created for the validation corpus. 
Recall that the manual annotations were not used to
train the labelling process.
As described in Section \ref{sec:wordlist} we used both ROC and F-1 measures to evaluate the performance of the classification.
Figure \ref{fig:maxdb-unsup-results} shows our ROC results for PostgreSQL, MaxDB and MySQL. We omit \textsf{exp1} due to poor results. We describe F-1 results in the text below.

@@

Because our ground truth annotations were relevant only to ISO9126,
 \textsf{exp1} had poor
performance due to the overlap between ISO9126 and the Kayed ontology (i.e., we annotated topics with labels which did not appear in the validation corpus). 
For \textsf{exp1} the F-measures for the NFRs for MaxDB were from $0$ to $0.18$ with an average (of all NFRs)
of $0.03$, for MySQL were from $0$ to $0.16$ with an average of
$0.05$, and for PostgreSQL $0$ to $0.15$ with an average of $0.07$.

@@

\subsubsection{Analysis of the Supervised Labelling}
% \label{sec:suplabelling}
Because our data-set was of word counts we expected Bayesian techniques, often used in spam filtering, to perform well. 
We tried other learners that WEKA~\cite{weka09} provides: rule learners, decision tree learners, vector space learners, and support vector machines.  
Figure \ref{fig:best-learn-per-tag} shows the performance of the best
performing learner per label: 
% We considered the best learner for a label to be 
the learner that had the highest ROC value for that label. 
The best learner is important because one uses a single learner per
label. When applying our technique, for each NFR one should select the
best performing learner possible. Given the tractability of these techniques, a tool which applied all learners and presented the best result should be feasible. 
%Figure \ref{fig:best-learn-per-tag} uses the ZeroR learner as a baseline, a common practice in machine learning,
%as ZeroR naively chooses the largest category all of the time. 
%ZeroR occasionally outperforms other learners; for labels which are not as common, this is to be expected because any miscategorization will hurt accuracy. 
%This is why we use ROC values, instead of accuracy, as they can better represent performance on labels which are not applicable to the majority of samples.

Figure \ref{fig:best-learn-per-tag} shows that MaxDB and MySQL have
quite different results, as the ROC values for \emph{reliability} and
\emph{functionality} swap between projects. For PostgreSQL, the
performance is nearly always poorer than the other two systems. The
reason for this lack of performance could be that the number of
topics, $N$, could be non-optimal for PostgreSQL. Perhaps topics were
getting too mixed and it was becoming hard to distinguish one topic
from the next. PostgreSQL's data-set was the largest of the 3 projects:
the XML data-dump that described the PostgreSQL topics that we
annotated was 8 times larger than MySQL and 1.7 times larger than MaxDB. These
size differences arise from the number of commits, the number of files
and the verbosity of the commit descriptions.

@@

As noted in Section \ref{sec:wordlist}, each topic in our data-set can be composed of zero or more NFRs. 
For example, a commit message might address \textit{reliability} in the context of \textit{efficiency}, or make a \textit{maintainability} improvement
in the source code that related to \textit{usability}. 
However, traditional machine learning techniques, such as Naive Bayes, can only map topics to a single class. 
The Mulan~\cite{mulan} library encapsulates several different multi-label machine learners which can label elements with multiple labels.
Mulan also includes methods for determining the performance of such techniques.
% TOO DETAILED The problem framed in the learners above has changed; instead of looking at the precision and recall of applying one label, we rank
% multiple labels at once. We must check if the full subset of labels was applied, and then how much of that subset was applied.

@@

\emph{Construct validity} -- we used only commit messages rather than
mail or bug tracker messages. 
To extend further we would need matching repositories for each project.
%Each case study was BecauseSince we required a seamless record, it was not possible to obtain
%other data sources. 
Possibly they would have influenced our results, but there would be a degree of correlation between the corpora.
 It is possible for a given label to occur across the arbitrary 30-day boundary we set. We suspect but have not proved that this is insignificant. 
%Conversely, construct validity was addressed in our validation as we used the files, topics and revisions to annotate the commit messages. 
%We used mailing-lists to verify the purpose behind some events.
Our taxonomy for software NFRs is subject to dispute, but seems to be
generally accepted. Finally, there are exogenous sources, such as
in-person discussions, which we did not access.  %We did not show these results to the stakeholders.

@@

Our data and scripts are available at \url{http://softwareprocess.es/nomen/}

@@

% 
% \begin{acknowledgements}
% 	Thanks to the helpful comments from the external reviewers.
% \end{acknowledgements}
